Combinatorial Proofs and Decomposition Theorems for First-order Logic

### 2021-01-10 main.tex
graphs

\usepackage[utf8]{inputenc}
\usepackage{dashbox}
\usepackage{ebproof}
\usepackage{amsmath}
\usepackage{ebproof}
\usepackage{virginialake}

@@

\newcommand{\VAR}{\textsc{var}}
\newcommand{\FUN}{\textsc{fun}}
\newcommand{\PRED}{\textsc{pred}}
\newcommand{\ATOM}{\textsc{atom}}
\newcommand{\FORM}{\textsc{form}}
\newcommand{\TERM}{\textsc{term}}
% *** MISC UTILITY PACKAGES ***
%
%\usepackage{ifpdf}
% Heiko Oberdiek's ifpdf.sty is very useful if you need conditional
% compilation based on whether the output is pdf or dvi.
% usage:
% \ifpdf
%   % pdf code
% \else
%   % dvi code
% \fi
% The latest version of ifpdf.sty can be obtained from:
% http://www.ctan.org/pkg/ifpdf
% Also, note that IEEEtran.cls V1.7 and later provides a builtin
% \ifCLASSINFOpdf conditional that works the same way.
% When switching from latex to pdflatex and vice-versa, the compiler may
% have to be run twice to clear warning/error messages.

@@

\newcommand{\set}[1]{\{#1\}}
\newcommand{\sqn}[1]{\vdash#1}

\newcommand{\sublist}[1]{\set{#1}}
\newcommand{\subst}[2]{#1\mathord{\scriptstyle\mapsto} #2}
\newcommand{\ssubst}[2]{\sublist{\subst{#1}{#2}}}

@@

We start from a countable set $\VAR=\set{x,y,z,\ldots}$ of variables,
a countable set $\FUN=\set{f,g,\ldots}$ of function symbols, and a
counrable set $\PRED=\set{p,q,\ldots}$ of predicate symbols. Each function symbol and
each predicate symbol has a finite arity. The grammars below generate
the set $\TERM$ of \bfit{terms}, denoted by $s,t,u,\ldots$, the set
$\ATOM$ of \bfit{atoms}, denoted by $a,b,c,\ldots$, and the set
$\FORM$ of \bfit{formulas}, denoted by $A,B,C,\ldots$:
\begin{equation*}
  \begin{array}{r@{~}l}
    t &\coloneqq~ x \mid f(t_1,\dots,t_n)
    \\[.5ex]
    a &\coloneqq~ \ttt\mid\fff\mid p(t_1,\dots,t_n) \mid \dual p(t_1,\dots,t_n)
    \\[.5ex]
    A &\coloneqq~ a \mid A\vlan A \mid A\vlor A \mid \exists x.A \mid \forall x.A
\end{array}
\end{equation*}
where the arity of $f$ and $p$ is $n$. Note, that in this paper we
consider the truth constants $\ttt$ (\emph{true}) and $\fff$
(\emph{false}) as atoms, and we consider all formulas in negation
normal form. The \bfit{negation} $(\dual\cdot)$ is defined for all
atoms and formulas via the De Morgan laws as follows:
\begin{equation*}
  \begin{array}{cc}
  \dual\ttt =\fff &
  \dual{p(t_1,\ldots,t_n)} = \dual{p}(t_1,\ldots,t_n) \\[.5ex]
  \dual\fff = \ttt &
  \dual{\dual p(t_1,\ldots,t_n)} = p(t_1,\ldots,t_n) \\[.5ex]
  \dual{A \vlan B} = \dual{A} \ \vlor \ \dual{B} &
  \dual{\exists x.A} = \forall x.\dual{A} \\[.5ex]
  \dual{A \vlor B} = \dual{A} \ \vlan \ \dual{B} & 
  \dual{\forall x.A} = \exists x.\dual{A} 
  \end{array}
\end{equation*}

\todo{add missing equations and make it look nice}

@@

A \bfit{directed graph} $G = (V, E)$ consists of a finite set $V$ of
\bfit{vertices} and a set of $E \subseteq V \times V$ of \bfit{direct edges}. An
\bfit{undirected graph} $G = (V, E)$ consists of a finite set $V$ of vertices
and a set of $E$ of \bfit{undirected edges} (two-element subsets of $V$).
For a given (directed or undirected) graph $G$, we write $V_G$ and $E_G$ for its
vertex and edge sets respectively. We write $vw$ for an edge $(v, w)$ or $\{v,
w\}$. Let $G = (V, E)$ and $G' = (V', E')$ be two undirected graphs. Without
loss of generality, we assume that $V$ and $V'$ are disjoint. The \bfit{union}
$G + G'$ is the graph $(V \cup V', E \cup E')$ and the \bfit{join} $G \times G'$
is the graph $(V \cup V', E \cup E' \cup \{vv' | v \in V, v' \in V'\})$. A
\bfit{cograph} is a $P_4$-free undireced grpah, i.e., for any distinct vertices $v_1,
v_2, v_3, v_4 \in V$, the restriction of the edge sets on them is not equal to
$\{v_1v_2, v_2v_3, v_3v_4\}$. In a graph $G = (V, E)$, the \bfit{neighbourhood}
$N(v)$ of $v \in V$ is defined as the set $\{w | vw \in E\}$, a set $W \subseteq
V$ induces a \bfit{matching} if for all $w \in W$, $N(w) \cap E$ is a singleton, and a module is a
set $M \subseteq V$ such that $N(v) \ M = N(w) \ M$ for all $v, w \in M$. A
\bfit{module} $M$ is strong if for every module $M'$, we have $M' \subseteq M$, $M
\subseteq$ or $M \cap M' = \emptyset$. A \bfit{graph homomorphism} $f: (V, E)
\rightarrow (V', E')$ is a function $f : V \rightarrow V'$ such that $f(v)f(w)
\in E'$ whenever $vw \in E$.

@@

\subsection{Fographs}

>>>>

\usepackage{xifthen}
\usepackage[utf8]{inputenc}
\usepackage{dashbox}
\usepackage{ebproof}
\usepackage{amsmath}
\usepackage{ebproof}
\usepackage{virginialake}

@@

\newcommand{\VAR}{\textsc{var}}
\newcommand{\FUN}{\textsc{fun}}
\newcommand{\PRED}{\textsc{pred}}
\newcommand{\ATOM}{\textsc{atom}}
\newcommand{\FORM}{\textsc{form}}
\newcommand{\TERM}{\textsc{term}}

\newcommand{\fequ}{\equiv}
\newcommand{\fequp}{\mathrel{\equiv'\mkern-2mu}}

\newcommand{\graph}[1]{\mathcal{#1}}
\newcommand{\vertices}[1][]{\ifthenelse{\isempty{#1}}{V}{V_{\graph{#1}}}}
\newcommand{\edges}[1][]{\ifthenelse{\isempty{#1}}{E}{E_{\graph{#1}}}}

\newcommand{\gA}{\graph{A}}
\newcommand{\gC}{\graph{C}}
\newcommand{\gG}{\graph{G}}
\newcommand{\gH}{\graph{H}}

\newcommand{\vA}{\vertices[A]}
\newcommand{\vC}{\vertices[C]}
\newcommand{\vG}{\vertices[G]}
\newcommand{\vH}{\vertices[H]}

\newcommand{\eA}{\edges[A]}
\newcommand{\eC}{\edges[C]}
\newcommand{\eG}{\edges[G]}
\newcommand{\eH}{\edges[H]}

% *** MISC UTILITY PACKAGES ***
%
%\usepackage{ifpdf}
% Heiko Oberdiek's ifpdf.sty is very useful if you need conditional
% compilation based on whether the output is pdf or dvi.
% usage:
% \ifpdf
%   % pdf code
% \else
%   % dvi code
% \fi
% The latest version of ifpdf.sty can be obtained from:
% http://www.ctan.org/pkg/ifpdf
% Also, note that IEEEtran.cls V1.7 and later provides a builtin
% \ifCLASSINFOpdf conditional that works the same way.
% When switching from latex to pdflatex and vice-versa, the compiler may
% have to be run twice to clear warning/error messages.

@@

\newcommand{\tuple}[1]{\langle#1\rangle}
\newcommand{\pair}[1]{(#1)}
\newcommand{\set}[1]{\{#1\}}
\newcommand{\sqn}[1]{\vdash#1}

\newcommand{\compl}[1]{#1^\complement}
\newcommand{\sublist}[1]{\set{#1}}
\newcommand{\subst}[2]{#1\mathord{\scriptstyle\mapsto} #2}
\newcommand{\ssubst}[2]{\sublist{\subst{#1}{#2}}}

@@

We start from a countable set $\VAR=\set{x,y,z,\ldots}$ of variables,
a countable set $\FUN=\set{f,g,\ldots}$ of function symbols, and a
counrable set $\PRED=\set{p,q,\ldots}$ of predicate symbols. Each function symbol and
each predicate symbol has a finite arity. The grammars below generate
the set $\TERM$ of \bfit{terms}, denoted by $s,t,u,\ldots$, the set
$\ATOM$ of \bfit{atoms}, denoted by $a,b,c,\ldots$, and the set
$\FORM$ of \bfit{formulas}, denoted by $A,B,C,\ldots$:
\begin{equation*}
  \begin{array}{r@{~}l}
    t &\coloneqq~ x \mid f(t_1,\dots,t_n)
    \\[.5ex]
    a &\coloneqq~ \ttt\mid\fff\mid p(t_1,\dots,t_n) \mid \dual p(t_1,\dots,t_n)
    \\[.5ex]
    A &\coloneqq~ a \mid A\vlan A \mid A\vlor A \mid \exists x.A \mid \forall x.A
\end{array}
\end{equation*}
where the arity of $f$ and $p$ is $n$. Note, that in this paper we
consider the truth constants $\ttt$ (\emph{true}) and $\fff$
(\emph{false}) as atoms, and we consider all formulas in negation
normal form. The \bfit{negation} $(\dual\cdot)$ is defined for all
atoms and formulas via the De Morgan laws as follows:
\begin{equation*}
  \begin{array}{cc}
  \dual\ttt =\fff &
  \dual{p(t_1,\ldots,t_n)} = \dual{p}(t_1,\ldots,t_n) \\[.5ex]
  \dual\fff = \ttt &
  \dual{\dual p(t_1,\ldots,t_n)} = p(t_1,\ldots,t_n) \\[.5ex]
  \dual{A \vlan B} = \dual{A} \ \vlor \ \dual{B} &
  \dual{\exists x.A} = \forall x.\dual{A} \\[.5ex]
  \dual{A \vlor B} = \dual{A} \ \vlan \ \dual{B} & 
  \dual{\forall x.A} = \exists x.\dual{A} 
  \end{array}
\end{equation*}
\todo{add missing equations and make it look nice}
\lutz{don't we need   $\dual{\dual a} = a$? (then it should follow that $\dual{\dual A} = A$)}

@@

A \bfit{graph} $\gG=\tuple{\vG,\eG}$ is a pair where $\vG$ is a finite
set of \bfit{vertices} and $\eG$ is a finite set of \bfit{edges},
which are two-element subsets of $\vG$. We write $vw$ for an edge
$\set{v,w}$.

The \bfit{complement} of a graph $\gG=\tuple{\vG,\eG}$ is
the graph $\compl\gG=\tuple{\vG,\compl\eG}$ where $vw\in\compl\eG$ iff
$vw\neq\eG$.

Let $\gG=\tuple{\vG,\eG}$ and $\gH=\tuple{\vH,\eH}$ be graphs such
that $\vG\cap\vH=\emptyset$. A \bfit{homomorphism}
$\phi\colon\gG\to\gH$ is a function $\phi\colon\vG\to\vH$ such that if
$vw\in\eG$ then $\phi(v)\phi(w)\in\eH$. The \bfit{union} $\gG +\gH$ is
the graph $\tuple{\vG \cup\vH,\eG\cup \eH}$ and the \bfit{join} $\gG
\times \gH$ is the graph $\tuple{\vG \cup\vH,\eG \cup \eH \cup \set{vw
    | v \in \vG, w \in \vH}}$.  A graph $\gG$ is \bfit{disconnected}
if $\gG=\gG_1+\gG_2$ for two non-empty graphs $\gG_1,\gG_2$, otherwise
it is \bfit{connected}. It is \emph{coconnected} if its complement is
connected.

A graph $\gG$ is \bfit{labelled} in a set $L$ if
each vertex $v\in\vG$ has an element $\ell(v)\in L$ associated with it, its \bfit{label}.

A \bfit{directed graph} $\gG =\tuple{\vG,\eG}$ is set $V$ of
\bfit{vertices} and a set $\eG\subseteq \vG\times\vG$ of \bfit{direct edges}.
A \bfit{directed graph homomorphism}
$\phi\colon\tuple{\vG,\eG}\to\tuple{\vH,\eH}$ is function $\phi\colon\vG\to\vH$ such that if
$\pair{v,w}\in\eG$ then $\pair{\phi(v),\phi(w)}\in\eH$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

An
\bfit{undirected graph} $G = (V, E)$ consists of a finite set $V$ of vertices
and a set of $E$ of \bfit{undirected edges} (two-element subsets of $V$).
For a given (directed or undirected) graph $G$, we write $V_G$ and $E_G$ for its
vertex and edge sets respectively. We write $vw$ for an edge $(v, w)$ or $\{v,
w\}$. Let $G = (V, E)$ and $G' = (V', E')$ be two undirected graphs. Without
loss of generality, we assume that $V$ and $V'$ are disjoint. The \bfit{union}
$G + G'$ is the graph $(V \cup V', E \cup E')$ and the \bfit{join} $G \times G'$
is the graph $(V \cup V', E \cup E' \cup \{vv' | v \in V, v' \in V'\})$. A
\bfit{cograph} is a $P_4$-free undireced grpah, i.e., for any distinct vertices $v_1,
v_2, v_3, v_4 \in V$, the restriction of the edge sets on them is not equal to
$\{v_1v_2, v_2v_3, v_3v_4\}$. In a graph $G = (V, E)$, the \bfit{neighbourhood}
$N(v)$ of $v \in V$ is defined as the set $\{w | vw \in E\}$, a set $W \subseteq
V$ induces a \bfit{matching} if for all $w \in W$, $N(w) \cap E$ is a singleton, and a module is a
set $M \subseteq V$ such that $N(v) \ M = N(w) \ M$ for all $v, w \in M$. A
\bfit{module} $M$ is strong if for every module $M'$, we have $M' \subseteq M$, $M
\subseteq$ or $M \cap M' = \emptyset$. A \bfit{graph homomorphism} $f: (V, E)
\rightarrow (V', E')$ is a function $f : V \rightarrow V'$ such that $f(v)f(w)
\in E'$ whenever $vw \in E$.
\subsection{Fographs}
